---
title: "Linear Models"
author: "Juan Martinez"
date: "3/31/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(readr)
library(gdata)
library(rstan)
library(brms)
library(parallel)
library(bayestestR)
library(bayesplot)
library(rstanarm)
library(broom)
library(insight)
library(mlbench)
library(PerformanceAnalytics)
```

# Introduction

We will compare naive and bayesian regression models in public transportation demand:

* Poisson
* Negative Binomial
* Zero-Inflated 
* Hurdle

The naive version of these models is already trained, so we only train and test the bayesian version.

## Transit data

```{r}
pre_board_train <- read_csv('C:/Users/jcmar/OneDrive/Documents/Spring_2023/Statistical_Rethinking/Data/new_bus_stop_12/pre_lock_train_data.csv')
pre_board_train$hour <- as_factor(pre_board_train$hour)
pre_board_train$month <- as_factor(pre_board_train$month)
pre_board_train$service_kind <- as_factor(pre_board_train$service_kind)

pre_board_test <- read_csv('C:/Users/jcmar/OneDrive/Documents/Spring_2023/Statistical_Rethinking/Data/new_bus_stop_12/pre_lock_test_data.csv')
```

The `pre_board_train` and `pre_board_test` datasets have been upload, so we proceed to standardize the variables before defining the prior distributions for each explanatory variable.

### `mean_temp`

The overall mean temperature distribution is:

```{r}

pre_board_train %>%
  ggplot(aes(x = mean_temp)) +
  geom_histogram()

```

Mean temperature grouped by hour:

```{r}
pre_board_train %>%
  ggplot(aes(x = mean_temp)) +
  geom_histogram() +
  facet_wrap(~hour)
```
Then,

```{r}
mean_temp_model <- glm(board_count ~ mean_temp, family = poisson(), data = pre_board_train)
summary(mean_temp_model)
```

### `mean_precip`

Precipitation distribution:

```{r}

pre_board_train %>%
  ggplot(aes(x = mean_precip)) +
  geom_histogram()

```

Then,

```{r}
mean_temp_model <- glm(board_count ~ mean_temp, family = poisson(), data = pre_board_train)
summary(mean_temp_model)
```


### `month_average_board_count`

```{r}

pre_board_train %>%
  ggplot(aes(x = month_average_board_count)) +
  geom_histogram()

```

Then,

```{r}
mean_month_average_board_count_model <- glm(board_count ~ month_average_board_count,
                                            family = poisson(), data = pre_board_train)
summary(mean_month_average_board_count_model)
```

### `surrounding_board_count`


```{r}

pre_board_train %>%
  ggplot(aes(x = surrounding_board_count)) +
  geom_histogram()

```

Then,

```{r}
surrounding_board_count_model <- glm(board_count ~ surrounding_board_count,
                                     family = poisson(), data = pre_board_train)

summary(surrounding_board_count_model)
```

# Bayesian Model with hour of the day as predictor

Variable distribution:

```{r}
pre_board_train %>%
  group_by(hour) %>%
  count() %>%
  ggplot(aes(x = hour, y = n)) +
  geom_bar(stat = 'identity')
```

Prior distribution:


```{r}
prior_hour <- prior_string("normal(0, 1.5)", class = 'b', coef = paste('hour', 5:23, sep =''))
```


to train the poisson model:

```{r}

bayes_pre_poiss_model_hour <- brm(board_count ~ 1 + hour,
                                  data = pre_board_train,
                                  family = poisson(), cores = 4,
                                  prior = prior_hour)
```

Bootstrap method chapter 7

Summary:

```{r}
summary(bayes_pre_poiss_model_hour)
```

Plots:

```{r}
plot(bayes_pre_poiss_model_hour)
```

```{r}
describe_prior(bayes_pre_poiss_model_hour)
```



```{r}
describe_posterior(bayes_pre_poiss_model_hour)
```


Leave-one-out cross-validation:

```{r}
loo(bayes_pre_poiss_model_hour, cores = getOption("mc.cores", 1))
```

## Predictions:

```{r}
bayes_pre_poiss_model_hour_pred <- predict(bayes_pre_poiss_model_hour, pre_board_test)
bayes_pre_poiss_model_hour_pred <- as.data.frame(bayes_pre_poiss_model_hour_pred)

bayes_pre_poiss_model_hour_test_RMSE <- sqrt(mean((bayes_pre_poiss_model_hour_pred$Estimate - pre_board_test$board_count)^{2}))

bayes_pre_poiss_model_hour_test_RMSE
```

Plots:

```{r}
bayes_pre_poiss_model_hour_pred %>%
  ggplot(aes(x = pre_board_test$board_count, y = bayes_poiss_model_pred$Estimate)) +
  geom_point(alpha = 0.25) +
  scale_y_continuous(breaks=c(0:10)) +
  scale_x_continuous(breaks=c(0:10))
```
```{r fig.width=20}
bayes_pre_poiss_model_hour_pred %>%
  ggplot(aes(x = (1:dim(bayes_pre_poiss_model_hour_pred)[1]), y = Estimate)) +
  geom_point(aes(colour = 'estimated'), alpha = 0.25) +
  geom_line(aes(colour = 'estimated'), alpha = 0.25) +
  geom_ribbon(aes(ymin = Q2.5, ymax = Q97.5, fill = 'estimated'), alpha = 0.2) +
  geom_point(aes(x = (1:dim(bayes_pre_poiss_model_hour_pred)[1]), y = pre_board_test$board_count, colour = 'obs'),
             alpha = 0.25) +
  geom_line(aes(x = (1:dim(bayes_pre_poiss_model_hour_pred)[1]), y = pre_board_test$board_count, colour = 'obs'),
             alpha = 0.25) 
```


# Bayesian Model with Hourly Mean Temperature as predictor

My first approach only considers a prior distribution for both the coefficient and intercept:


```{r}
prior_mean_temp <- prior("normal(0, 0.393)", class = 'b', coef = 'mean_temp')
```

Then,

```{r}
bayes_pre_poiss_model_mean_temp <- brm(board_count ~ 1 + mean_temp,
                                       data = pre_board_train,
                                       family = poisson(), cores = 4,
                                       prior = prior_mean_temp)

summary(bayes_pre_poiss_model_mean_temp)
```

```{r}
plot(bayes_pre_poiss_model_mean_temp)
```


The second approach considers a unique prior distribution for the coefficient and intercept:


```{r}
prior_mean_temp <- c(prior("normal(0, 0.393)", class = 'b', coef = 'mean_temp'),
                     prior("normal(0, 0.482)", class = 'Intercept'))
```

Then,

```{r}
bayes_pre_poiss_model_mean_temp <- brm(board_count ~ 1 + mean_temp,
                                       data = pre_board_train,
                                       family = poisson(), cores = 4,
                                       prior = prior_mean_temp)

summary(bayes_pre_poiss_model_mean_temp)
```

```{r}
plot(bayes_pre_poiss_model_mean_temp)
```

## Predictions

```{r}
bayes_pre_poiss_model_mean_temp_pred <- predict(bayes_pre_poiss_model_mean_temp,
                                                pre_board_test)

r.sq <- as.character(round(summary(lm(pre_board_test$board_count ~ bayes_pre_poiss_model_mean_temp_pred[,1]))$r.squared, 2))

r.sq2 <- paste('R^2 == ', r.sq)

ggplot() +
  geom_point(aes(x = bayes_pre_poiss_model_mean_temp_pred[,1],
                 y = pre_board_test$board_count)) +
  geom_errorbar(aes(x = bayes_pre_poiss_model_mean_temp_pred[,1],
                    y = pre_board_test$board_count,
                    xmin = bayes_pre_poiss_model_mean_temp_pred[,1] -
                      bayes_pre_poiss_model_mean_temp_pred[,2],
                    xmax = bayes_pre_poiss_model_mean_temp_pred[,1] +
                      bayes_pre_poiss_model_mean_temp_pred[,2])) +
  geom_smooth(aes(x = bayes_pre_poiss_model_mean_temp_pred[,1],
                  y = pre_board_test$board_count), method = 'lm',
              color = "red", lty = 2) +
  geom_text(aes(x= 6, y=9.5, label = r.sq2, size = 8), parse=TRUE, show.legend = F) +
  xlab("Predicted") +
  ylab("Observed")

```


# Full Bayesian Poisson Regression Model

```{r}
priors <- c(prior_string("normal(0, 1.5)", class = 'b', coef = paste('hour', 5:23, sep ='')),
            prior("normal(0, 0.393)", class = 'b', coef = 'mean_temp'),
                     prior("normal(0, 0.482)", class = 'Intercept'))
```

Then,

```{r}
bayes_pre_poiss_model <- brm(board_count ~ 1 + hour + mean_temp,
                             data = pre_board_train,
                             family = poisson(), cores = 4,
                             prior = priors)

summary(bayes_pre_poiss_model)
```


```{r}
plot(bayes_pre_poiss_model)
```


# Naive Poisson Regression Model

Comparison with the Naive Poisson model:

```{r}
Pre_Poisson_model <- readRDS('C:/Users/jcmar/OneDrive/Documents/Spring_2023/Statistical_Rethinking/Data/bus_stop_12/Pre_Poisson_model.rds')
summary(Pre_Poisson_model)

```

# Complete Bayesian Model

```{r}
priors_var = c(prior(normal(0, 0.9), class = b, coef = 'month'),
               prior(normal(0, 0.95), class = Intercept, coef = 'month'),
               prior(normal(0, 1.1), class = b, coef = 'service_kind'),
               prior(normal(0, 1.15), class = Intercept, coef = 'service_kind'),
               prior(normal(0, 1), class = b, coef = 'hour'),
               prior(normal(0, 1.5), class = Intercept, coef = 'hour'),
               prior(normal(0, 2.789), class = b, coef = 'mean_temp'),
               prior(normal(0, 0.17), class = Intercept, coef = 'mean_temp'),
               prior(normal(0, 0.214), class = b, coef = 'mean_precip'),
               prior(normal(0, 0.544), class = Intercept, coef = 'mean_precip'),
               prior(normal(0, 0.010), class = b, coef = 'month_average_board_count'),
               prior(normal(0, 0.569), class = Intercept, coef = 'month_average_board_count'),
               prior(normal(0, 0.367), class = b, coef = 'surrounding_board_count'),
               prior(normal(0, 0.506), class = Intercept, coef = 'surrounding_board_count'))

bayes_pre_poiss_model2 <- brm(board_count ~ .,
                              data = pre_board_train,
                              family = poisson(), cores = 4,
                              prior = priors_var)
```



```{r}
priors_by_fn <- get_prior(board_count ~ .,
                          data = pre_board_train,
                          family = poisson())

priors_by_fn
```

